### 2021-11-11 NSF2021/Description/specificaim2.tex
working on SA2

In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
%\begin{wrapfigure}[13]{l}{1.8in}
%  \centering
%  \input{figures/doublewell.tikz}
%  \caption{\label{fig:well} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
%\end{wrapfigure}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha (dF^N)^{-1} F(u^N),
\end{align*}
where $dF^N = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative
of $F$, and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest computational challenge is
solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
\begin{wrapfigure}[]{l}{1.8in}
  \vspace{-46pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} (a) The double well potential $U_1 =
  0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
  and inexact Newton's methods.}
\end{wrapfigure}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging since the fundamental solution is not easily
determined. We propose to instead use an inexact Newton's method where
$dF^{-1}$ is defined in terms of the piecewise quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering a finite difference method applied to its
one-dimensional version in $(0,1)$ with the Dirichlet boundary
conditions $u(0) = U_0$ and $u(1) = U_1$.  These parameters result in
the double well potential illustrated in Figure~\ref{fig:CA}(a), and the
inexact Newton's method converges to the solution in
Figure~\ref{fig:CA}(b). Figure~\ref{fig:CA}(c) shows the max norm of the
residual at each Newton iterate when using the exact second derivative
(red curve) and the piecewise constant approximate (blue curve).

Solutions of the screened Poisson equation can be written
in terms of volume and layer potentials. We first define the level set
$\Gamma = \{\xx \:|\: \uu(\xx) = (U_0 + U_1)/2 \}$ which separates the
domain $\Omega$ into two regions. Then, assuming that the particles are
all contained in the hydrophobic region, a solution of the screened
Laplace equation (i.e.~equation~\eqref{eq:screenedPoisson} when $F[u^N]
= 0$) is
\begin{align*}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nn_\yy} K_0 \left(
    \frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) ds_\yy + 
    \frac{1}{2\pi} \int_{\Gamma} K_0 \left(
    \frac{|\xx - \yy|}{\rho} \right) \eta(\yy) ds_\yy,
\end{align*}

>>>>

In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  J[u] = -\rho^2 \Delta u + f'(u).
\end{align}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha (dJ^N)^{-1} F(u^N),
\end{align*}
where $dJ^N = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative
of $F$, and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest computational challenge is
solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = J[u^N].
\end{align}
\begin{wrapfigure}[]{l}{1.8in}
  \vspace{-46pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} (a) The double well potential $u_0 =
  0.5$, $u_1 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
  and inexact Newton's methods.}
\end{wrapfigure}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging since the fundamental solution is not easily
determined. We propose to instead use an inexact Newton's method where
$dJ^{-1}$ is defined in terms of the piecewise quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - u_0)^2, &\mbox{if } u \leq (u_0 + u_1)/2 \\
    (u - u_1)^2, &\mbox{if } u > (u_0 + u_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = J[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of exact and inexact Newton's method
applied to~\eqref{eq:F} by considering a finite difference method
applied to its one-dimensional version in $(0,1)$ with the Dirichlet
boundary conditions $u(0) = u_0$ and $u(1) = u_1$. These parameters
result in the double well potential illustrated in
Figure~\ref{fig:CA}(a), and both the exact and inexact Newton's methods
converge to the solution in Figure~\ref{fig:CA}(b).
Figure~\ref{fig:CA}(c) shows the max norm of the residual at each Newton
iterate when using the exact second derivative (red curve) and the
piecewise constant approximate (blue curve).

Having demonstrated that the inexact Newton method converges in one
dimension, it will be used to solve the Allen-Cahn equation in two more
or more dimensions. Since we are interested in solving~\eqref{eq:SL} in
general unbounded geometries, we will develop an integral equation
formulation. We first partition $\Omega$ into $\Omega_0 = \{\xx \in
\Omega \:|\: u(\xx) < (u_0 + u_1)/2\}$ and $\Omega_1 = \{\xx \in \Omega
\:|\: u(\xx) > (u_0 + u_1)/2\}$, and we let $\Lambda = \{\xx \:|\:
u(\xx) = (u_0 + u_1)/2 \}$ be their common boundary points. We then let
$G_0$ be the fundamental solution of $(-\rho^2 \Delta - 2a)$ and $G_1$
be the fundamental solution of $(-\rho^2 \Delta + 2)$. Then, the
solution of~\eqref{eq:screenedPoisson} can be written as $u = u^P + u^H$
where $u^P$ is a volume potential and $u^H$ is a layer potential. The
volume potential is
\begin{align*}
  u^{P}(\xx) = \int_{\Omega_i} G_i(\xx,\yy) J[u^N](\yy)\, d\yy, 
    \quad i = 1,2,
\end{align*}
but it is discontinuous across $\Lambda$ and does not satisfy the
prescribed hydrophobic boundary conditions on the particles $\Gamma_i$.
The boundary conditions are enforced with the layer potential
\begin{align*}
  u^{H}(\xx) = \begin{cases}
    \displaystyle\int_{\Lambda} G_0(\xx,\yy) \eta_0(\yy) \, ds_\yy + 
    \displaystyle\int_{\Gamma} \pderiv{}{\nn_\yy} G_0(\xx,\yy)
      \sigma(\yy) \,ds_\yy, \quad & \xx \in \Omega_0, \\
    \displaystyle\int_{\Lambda} G_0(\xx,\yy) \eta_1(\yy) \, ds_\yy,
     & \xx \in \Omega_1.
  \end{cases}
\end{align*}

Then, assuming that the particles are all contained in the
hydrophobic region, a solution of the screened Laplace equation
(i.e.~equation~\eqref{eq:screenedPoisson} when $G[u^N] = 0$) is
\begin{align*}
  u(\xx) = \frac{1}{2\pi} \int_{\bd\Omega} \pderiv{}{\nn_\yy} K_0 \left(
    \frac{|\xx - \yy|}{\rho}\right) \sigma(\yy) ds_\yy + 
    \frac{1}{2\pi} \int_{\Gamma} K_0 \left(
    \frac{|\xx - \yy|}{\rho} \right) \eta(\yy) ds_\yy,
\end{align*}
